---
title: "Week 03 Assignment 02_datasets"
author: "Josh Pickel"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    template: ./../latex-templates/homework.tex
  html_document:
    df_print: paged
  word_document: default
course: STATS 419 Survey of Multivariate Analysis
course-short: STATS419
instructor: Monte J. Shaffer
email: 11594029.joshua.pickel@wsu.edu
fig_caption: yes
number_sections: yes
---


```{r setup, message = FALSE, warning =FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message = FALSE}
library(devtools)
github.path = 'https://raw.githubusercontent.com/JoshuaPickel/WSU_STATS419_FALL2020/';
source_url(paste0(github.path,"master/functions/libraries.R"));
source_url( paste0(github.path,"master/functions/functions-imdb.R"));
```

# Matrix
Create the "rotate matrix" functions as described in lectures.  Apply to the example "myMatrix".

```{r, Matrix Rotations, message = FALSE}
source_url(paste0(github.path,"master/functions/functions-matrix.R"))
# Create myMatrix
myMatrix = matrix ( c (
1, 0, 2,
0, 3, 0,
4, 0, 5
), nrow=3,
byrow=T);
```

## Matrix 90 Degrees
```{r}
# Rotating clockwise.
rotateMatrix90(myMatrix)
```

## Matrix 180 Degrees
```{r}
# Rotating Clockwise
rotateMatrix180(myMatrix)
```

## Matrix 270 Degrees
```{r}
# Rotating Clockwise
rotateMatrix270(myMatrix)
```

# IRIS Plot
 Recreate the graphic for the IRIS Data Set using R. Same titles, same scales, same colors.
\newpage
## Plot
```{r, warning = FALSE,fig.cap = 'Plot From Plotting IRIS Data',fig.sum = c(6,6)}
library(datasets)
plot(iris[1:4],pch = 21,cex = 1,col = 'black', bg= c("red", "green3",
"blue")[iris$Species],main = 'Iris Data(red=sestosa,green=versicolor,blue=virginica)',gap.axis = 1.5)

```

# IRIS Summary
Right 2-3 sentences concisely defining the IRIS Data Set. Maybe search KAGGLE for
a nice template. Be certain the final writeup are your own sentences (make certain
you modify what you find, make it your own, but also cite where you got your ideas
from). NOTE: Watch the video, Figure 8 has a +5 EASTER EGG.

## Response
The Iris data set was created in 1936 by Edgar Anderson and is a multivariate data set
that containing four different measurements of three different Iris flower species.
The data set contains 50 records for each of the different flower species; Setosa,
Virginica, and Versicolor recording measurements for petal length, petal width, sepal
width and sepal length, all recorded in centimeters.

# Cleaning Personality Data
 Import "personality-raw.txt" into R. Remove the V00 column. Create two new
columns from the current column "date_test": year and week. Stack Overflow may help:
https://stackoverflow.com/questions/22439540/how-to-get-week-numbers-from-dates ...
Sort the new data frame by YEAR, WEEK so the newest tests are first ... The newest
tests (e.g., 2020 or 2019) are at the top of the data frame. Then remove duplicates
using the unique function based on the column "md5_email". Save the data frame in the
same "pipe-delimited format" ( | is a pipe ) with the headers. You will keep the new
data frame as "personality-clean.txt" for future work (you will not upload it at this
time). In the homework, for this tasks, report how many records your raw dataset had
and how many records your clean dataset has.

```{r,clean data, message = FALSE }

source_url(paste0(github.path,"master/functions/functions-cleanup.R"))
files = paste0(github.path,"master/datasets/personality-raw.txt")
personality_raw = read.csv(files,header = T, sep = '|')
personality_clean = removeColumn(personality_raw,'V00')
personality_clean = convertDates(personality_clean,personality_clean$date_test)
personality_clean = removeDuplicates(personality_clean,personality_clean$md5_email)
head(personality_clean)
```
## Raw Dataset Dimensions
```{r}
dim(personality_raw)
```

## Clean Dataset Dimensions
```{r}
dim(personality_clean)
```


# Custom Functions
Write functions for doSummary and sampleVariance and doMode ... test these
functions in your homework on the "monte.shaffer@gmail.com" record from the clean
dataset. Report your findings. For this "monte.shaffer@gmail.com" record, also
create z-scores. Plot(x,y) where x is the raw scores for "monte.shaffer@gmail.com"
and y is the z-scores from those raw scores. Include the plot in your assignment, and
write 2 sentences describing what pattern you are seeing and why this pattern is
present.

## Setup and Get Row For "monte.shaffer@gmail.com"
``` {r, message = FALSE}
source_url(paste0(github.path,"master/functions/functions-custom.R"))
monte_row = personality_clean[which(personality_clean$md5_email ==
'b62c73cdaf59e0a13de495b84030734e'),]

```
## doSummary
```{r ,doSummary}
doSummary(monte_row)
```
## custom_mode
```{r, custom_mode}
custom_mode(monte_row)
```
## doSampleVariance Naive
```{r,doSampleVariance Naive}
doSampleVariance(monte_row, 'naive')
```
## doSampleVariance Two Pass
```{r, doSampleVariance Two Pass}
doSampleVariance(monte_row,'Two Pass')
```

## Z-scores
```{r, zscores, fig.cap = 'Plotting Raw Scores Against Z-scores' }
zscores = zScores(monte_row)
plot(zscores$raw.scores,zscores$z.score,main = 'Raw Scores Vs Z-Scores',xlab = 'raw scores',ylab = 'z-scores')
```


## Response
After running the doSummary() on the record for ‘monte.shaffer@gmail.com’, it is
apparent that the mode is 4.2, the mean is 3.48, there are no NAN values, the length
is 64 columns, the two pass and naïve variances are both .7528136 and the standard
deviation is .8676483.
Considering the plot created from plotting raw scores against z-scores we can conclude
from Figure 9 that there is a strong positive linear relationship between raw scores
and z-scores. This is because the z-score is calculated from the raw values, meaning
the two are highly correlated. 

# Will vs Denzel
Compare Will Smith and Denzel Washington. You will have to create a new variable \$millions.2000 that converts each movie's \$millions based on the $year of the 
movie, so all dollars are in the same time frame. You will need inflation data from about 1980-2020 to make this work.

## Get Data For Actors
```{r,message = FALSE}
source_url(paste0(github.path,"master/functions/functions-imdb.R"))

# Add mill.2000 column to Will's data frame with converted dollar values
nmid = "nm0000226"
will = grabFilmsForPerson(nmid)

nmid = "nm0000243"
denzel = grabFilmsForPerson(nmid)
```

## Get Inflation Data
```{r, Inflation Data, warning = FALSE}
source_url(paste0(github.path,"master/functions/functions-inflation.R"))
inflation_data = grabInflation()
```

## Adjust Denzel and Will Million Values
```{r, adjusting Denzel and Will}

will = convertDollars(will,inflation_data)
denzel = convertDollars(denzel,inflation_data)

head(will$movies.50)
head(denzel$movies.50)

```
## Response
With converting all movie revenues to year 2000 dollars, we can tell that most of the
movies Will Smith is in have higher revenue than the movies featuring Denzel
Washington. We can also tell that Will Smith was in a movie that generated $336.03
million, which is considered an outlier. In general, we can see that the median
revenue is the almost same between Will Smith and Denzel Washington, and Will Smith
has a higher maximum revenue value than Denzel Washington. 

# Will Vs Denzel
Build side-by-side box plots on several of the variables (including #6) to compare the
two movie stars. After each box plot, write 2+ sentence describing what you are
seeing, and what conclusions you can logically make. You will need to review what the
box plot is showing with the box portion, the divider in the box, and the whiskers.

## Comparing Will vs Denzel Year 2000 Millions

```{r, Will vs Denzel} 
boxplot(denzel$movies.50$mill.2000,will$movies.50$mill.2000,main = 'Will vs Denzel Year 2000 Revenue (in millions)', names = c('Denzel Washington', 'Will Smith'), xlab = 'Actor' , ylab = 'Revenue(Millions)')
```

### Response
Building a boxplot comparing the revenues (scaled to year 2000) between Will Smith and
Denzel Washington (Figure 10), we can tell that Will Smith has a slightly higher
median revenue than Denzel Washington. We can also conclude that fifty percent of Will
Smith’s revenue values are between $30.45 million and $137.66 million and fifty
percent of Denzel Washington’s revenue values are between $22.39 million and $70.06
million. This shows us that Will Smith’s revenue values are spread out more when
compared to Denzel Washington’s’ values.

## Comparing Will vs Denzel Years
```{r, Will vs Denzel Years}
boxplot(denzel$movies.50$year,will$movies.50$year,main = 'Will vs Denzel Year', names = c('Denzel Washington', 'Will Smith'), xlab = 'Actor' , ylab = 'Year')

```

### Response
Comparing Will Smith and Denzel Washington’s years using a boxplot (Figure 11), there
are serval key findings. We can conclude that, for this data set, Denzel Washington
appears in older movies when compared to Will Smith. We can also determine that Denzel
Washington’s most recent film was before 2020, while Will Smith’s most recent film was
in 2020. We can also conclude that half of the movies Will Smith has been featured in
were after 2006, and the other half before 2006. Likewise, we can determine that half
of the movies Denzel Washington has appeared in were after 1999, and the other half
before 1999.


## Comparing Will vs Denzel Runtime
```{r, Will vs Denzel Runtime}
boxplot(denzel$movies.50$minutes,will$movies.50$minutes,main = 'Will vs Denzel Runtime', names = c('Denzel Washington', 'Will Smith'), xlab = 'Actor' , ylab = 'Runtime')
```

### Response
Using a boxplot to compare Will Smith and Denzel Washington’s movie
runtime(Figure 12), we can conclude that Denzel Washington has appeared in
the longest movie in the data set. We can also conclude that Will Smith
appeared in the shortest movie in the data set. Furthermore, we can conclude
that the upper half of Denzel Washington’s runtime values are higher than
Will Smith’s runtime values.
